---
title: "Forecasting completion dates using data from Jira and R"
author: "Sam"
date: 2020-04-01T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# The problem

I recently found myself in a position where I needed to forecast when something our team was working on would be ready.  Until that point we had quite happily been busy building without estimating the size of the tasks we were getting through.  I was keen to not have to start introducing an arbitrary system of estimations - or rely on any _gut feelings_.  By this point we had a been working together for some time, so we had enough data to be able to help with the forecasting.  

I decided to try to estimate the number of tasks per day we could get through.  I also wanted to have confidence intervals so we could communicate our forecast with more certainly.  In order to achieve this I bootstrapped the mean number of tickets that were moved on a given day to "done".  I then used the results of the bootstrap to be able to forecast different scenarios with various confidence intervals.

In this post, I will explain a bit how we created our forecast, using some tools from R and some basic data from Jira.


# Setup

I unfortunately didn't have access to the Jira API, so I had to make do with the CSV download.  I filtered for all the tickets that my team had worked on, then downloaded the file.
<Jira download>
 

This forecast was produced using the following packages.  We need dplyr for data manipulation and the Infer package for the statisitical inference.  I also used Janitor and Lubridate to help clean up the messy data you get by default when using the CSV download in Jira.

```{r eval = FALSE}
library(dplyr)
library(infer)
library(lubridate)
library(janitor)
```

# Data preparation

Before we can create our forecast we need to clean up our data a bit.  In this case the main cleaning to be done is setting dates to the correct class and removing spaces from column names.  

```{r pie, fig.cap='A fancy pie chart.', tidy=FALSE}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```
